---
title: "PA1_template.Rmd"
author: "Jad Melki"
date: "June 14, 2015"
output: html_document
---
# Loading and Processing the Data 

1. Read the data frame. 
``` {r} 
raw.data <- read.csv("activity.csv")
```

2. Use dplyr and plyr packages for efficient filtering and displaying.
``` {r}
library(dplyr)
library(plyr)
data <- tbl_df(raw.data)
```

# What is the mean total number of steps taken per day?

1. Produce a cross-tabulation that calculates the total number of steps taken per day. 
``` {r} 
xtabs(steps ~ date, data=data)
```

2. Make a histogram of the total number of steps taken each day
``` {r} 
hist(xtabs(steps ~ date, data=data))
```

3. Calculate mean number of steps taken per day
``` {r}
with(data, mean(steps, na.rm=TRUE))
```

4. Calculate median number of steps taken per day
``` {r}
median(as.vector((xtabs(steps ~ date, data=data))))
```

# What is the average daily activity pattern?

1. Make a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
data.intervals <- ddply(data, .(interval), summarise, mean(steps, na.rm=TRUE))
plot(data.intervals, type="l")
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
data.intervals.max <- filter(data.intervals, data.intervals[,2]==max(data.intervals[,2]))
as.numeric(data.intervals.max$interval)
```

# Imputing missing values

1. Calculate and report the total number of missing values in the dataset
```{r}
with(data, sum(is.na(steps)))
```

2. Devise a strategy for filling in all of the missing values in the dataset. 
My strategy is to use the mean for that day to replace the NAs of that day. 

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

I first create a cross-tabulation of the means for each day. Then copy the original data frame "data" into "data1". The create a loop that will replace the NA for each day with the mean steps for that day. 

```{r}
mean.steps <- xtabs(steps~date,aggregate(steps~date,data,mean))
data1 <- data
for (i in 1:61) 
  {
  filter(data1, date==names(mean.steps)[i])
  data1[is.na(data1)]<-mean.steps[i]
  }
```

4a. Make a histogram of the total number of steps taken each day. Do these values differ from the estimates from the first part of the assignment?
Both plots seem identical with no change occuring from the imputation of the NAs.
```{r}
par(mfcol= c(2,1))
hist(xtabs(steps ~ date, data=data))
hist(xtabs(steps ~ date, data=data1))
```

4b. Calculate and report the mean total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment?
There is a 5.1 difference between the means before and after imputation.

The mean number of steps taken per day before imputation is: 
``` {r}
with(data, mean(steps, na.rm=TRUE))
```
The mean number of steps taken per day after imputation is: 
``` {r}
with(data1, mean(steps, na.rm=TRUE))
```

4c. Calculate and report the median total number of steps taken per day. Do these values differ from the estimates from the first part of the assignment?
There is no difference in between the medians before and after imputation.

The median number of steps taken per day before imputation is: 
``` {r}
median(as.vector((xtabs(steps ~ date, data=data))))
```

The median number of steps taken per day after imputation is: 
``` {r}
median(as.vector((xtabs(steps ~ date, data=data1))))
```

What is the impact of imputing missing data on the estimates of the total daily number of steps?
There is a small impact on the mean steps per day but not on the median nor on the daily total steps. 


# Are there differences in activity patterns between weekdays and weekends?
Unfortunately, I'm out of time for this one!

